---
title: "Chapter 5 Computer Exercises"
output: html_document
---

```{r setup, include=FALSE}
library(wooldridge)
library(car)
```

C1. **Question:** Use the data in WAGE1 for this exercise. 

i) Estimate the equation 

$$wage = \beta_0 + \beta_1 educ + \beta_2 exper + \beta_3 tenure + u$$

Save the residuals and plot a histogram. 

```{r}
data("wage1")
model <- lm(wage ~ educ + exper + tenure, data = wage1)
residuals <- resid(model)
hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")
```

ii) Repeat i) but with log(wage) as the dependent variable

```{r}
model_log <- lm(log(wage) ~ educ + exper + tenure, data = wage1)
residuals_log <- resid(model_log)
hist(residuals_log, main = "Histogram of Residuals (Log Wage)", xlab = "Residuals")
``` 

iii) Would you say that Assumption MLR.6 is closer to being satiisfied for the level-level model or the log-level model?

Log-level because looks more normally distributed. 

C2. **Question:** Use the data in GPA2 for this exercise. 

i) Using all 4,137 obserations, estimate the equation

$$colGPA = \beta_0 + \beta_1 hsperc + \beta_2 sat + u$$

```{r}
data("gpa2")
model_gpa <- lm(colgpa ~ hsperc + sat, data = gpa2)
summary(model_gpa)
```

ii) Restimate using the first 2,070 obseratoins 
```{r}
model_gpa_half <- lm(colgpa ~ hsperc + sat, data = gpa2[1:2070, ])
summary(model_gpa_half)
```


iii) Find the ratio of the standard errors on $hsperc$ from parts i) and ii). Compare this with the result from 5.10

The ratio of standard errors can be calculated as follows:
```{r}
se_full <- summary(model_gpa)$coefficients["hsperc", "Std. Error"]
se_half <- summary(model_gpa_half)$coefficients["hsperc", "Std. Error"]
ratio_se <- se_full / se_half
ratio_se
```

```{r}
sqrt(2070/4137)
```


C3. **Question:** In equation (4.42) of Chapter 4, using the data set BWGHT, copmute the LM statistic for testing whether $motheduc$ and $fatheduc$ are jointly 
significant. In obtaining the residuals for the restricted model, be sure that the restricted moel is estimated 
using only those oservations for which all variables in the unrestricted model are available (see Example 4.9). 

```{r}
data("bwght")

# Filter for complete cases
bwght_complete <- bwght[complete.cases(bwght[, c("bwght", "cigs", "parity", "faminc", "motheduc", "fatheduc")]), ]

# Unrestricted model
unrestricted_model <- lm(bwght ~ cigs + parity + faminc + motheduc + fatheduc, data = bwght_complete)
# Restricted model
restricted_model <- lm(bwght ~ cigs + parity + faminc, data = bwght_complete)
# LM statistic calculation
residuals_restricted <- resid(restricted_model)

# Regress restricted residuals on all independent variables
auxiliary_reg <- lm(residuals_restricted ~ cigs + parity + faminc + motheduc + fatheduc, data = bwght_complete)
summary(auxiliary_reg)
r_squared <- summary(auxiliary_reg)$r.squared
n <- nrow(bwght_complete)
n* r_squared

# another way to do it, per claude
SSR_restricted <- sum(residuals_restricted^2)
SSR_unrestricted <- sum(resid(unrestricted_model)^2)
n <- nrow(bwght_complete)
k <- 2 # number of restrictions (motheduc and fatheduc)
LM_statistic <- (SSR_restricted - SSR_unrestricted) / SSR_unrestricted * n
LM_statistic
```


C4. **Question:** Detect nonnormality in underlying population distrbiutions. 

i) Use 401ksubs, keeping only observations with fsize = 1. Find the skewness measure for inc. Same for log(inc). Which has mroe skewness and seems less liekly to e normally distributed?
```{r}
data("k401ksubs")
install.packages("moments")
library(moments)
data_401k <- subset(k401ksubs, fsize == 1)
skewness_inc <- skewness(data_401k$inc, na.rm = TRUE)
skewness_log_inc <- skewness(log(data_401k$inc), na.rm = TRUE)
skewness_inc
skewness_log_inc
```

log(inc) seems mroe normally distributed than the inc. 

ii) Use BWGHT2. Find the skewness measures for beght and log(bwght). What do you conclude?
```{r}
data("bwght2")
skewness_bwght <- skewness(bwght2$bwght, na.rm = TRUE)
skewness_log_bwght <- skewness(log(bwght2$bwght), na.rm = TRUE)
skewness_bwght
skewness_log_bwght
```

bwght less skewed than log(bwght). This makes sense to me because birth weights cannot be negative or 0, so log transformation would skew right.


iii) The two examples above are contradictions to this statement; enough said. 

iv) If we are interested in the normality assumption in the context of regression, should we be evaluatiing the unconditional distributions of y and log(y)? Explain

No. We should be evaluating the conditional distributions of the error term u and log(u) given X.

C5. **Question:** Consider C11 analysis using HTV. 
i) How many different values are taken on by educ in the sample? Does educ have a continuous distribution?
```{r}
data("htv")
num_educ_values <- length(unique(htv$educ))
num_educ_values
unique(htv$educ)
```

no it is not continuous; 15 different values

ii) Plot a histogram of educ with a normal distribution overlay. Does the distribution of educ appear anything close to normal?

```{r}
hist(htv$educ, probability = TRUE, main = "Histogram of educ with Normal Overlay", xlab = "educ")
x_vals <- seq(min(htv$educ), max(htv$educ), length.out = 100)
y_vals <- dnorm(x_vals, mean = mean(htv$educ), sd = sd(htv$educ))
lines(x_vals, y_vals, col = "red")
```

no these don't look normal whatsoever

iii) Which of the CLM assumptions seems clearly violated in the model? How does this violation change the statistical inference procedures carried out in Computer Exercise C11 in Chapter 4. 

error can't be normally distributed because educ is not continuous, which violates MLR.6. This means that the t-stats and F-stats used in C11 may not be valid, but that depends on sample size.





C6. **Question:** Use the data in ECONMATH to answer the question. 

i) Logically, what are the smallest and largest values that can be taken on by the variable score?

```{r}
data("econmath")
min_score <- min(econmath$score, na.rm = TRUE)
max_score <- max(econmath$score, na.rm = TRUE)
min_score
max_score
```

ii) Consider the linear model $score = \beta_0 + \beta_1 colgpa + \beta_2 actmath + \beta_3 acteng + u$. Why cannot Assumption MLR.6 hold for error term u? What consequences does this have for 
using the t-stat to test $H0: \beta_3 = 0$?

Because score is bounded between 0 and 100, the error term u cannot be normally distributed as it would imply that score could take on values outside this range. 
This violates MLR.6 which assumes that the error term is normally distributed. 

As a consequence, using the t-statistic to test $H0: \beta_3 = 0$ may not be valid since the underlying assumption of normality is violated. But this will depend on sample size.

iii) Estimate the model from part (ii) and obtain the t-stat and associated p-value for testing $H0: \beta_3 = 0$. How would you defend your findings to someone who makes the following statement: 
"You cannot trust that p-value because clearly the error term in the equation cannot have a normal distribution"

```{r}
data("econmath")
model_score <- lm(score ~ colgpa + actmth + acteng, data = econmath)
summary(model_score)

# t-statistic for acteng
beta_acteng <- summary(model_score)$coefficients["acteng", "Estimate"]
se_acteng <- summary(model_score)$coefficients["acteng", "Std. Error"]
t_stat_acteng <- beta_acteng / se_acteng
t_stat_acteng

nrow(econmath)
```
sample size is pretty big 856


